---
title: "Select cutoff for final clustering"
author: "Zhiyuan"
date: "21/10/2020 (last modified: `r Sys.Date()`)"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.width = 9, fig.keep = FALSE, fig.path = "plots/dendritic/cutoff_selection/Rmarkdown_")
knitr::opts_knit$set(root.dir = "~/OneDrive - Nexus365/Project ClinCluster/evaluate-integration/evaluate-integration")

library(Seurat)

library(limma)
library(edgeR)
library(mclust)

library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(igraph)
library(viridis)
library(pheatmap)

source("~/OneDrive - Nexus365/Project ClinCluster/evaluate-integration/evaluate-integration/functions_simulate.R")
source("~/OneDrive - Nexus365/Project ClinCluster/evaluate-integration/evaluate-integration/functions_pipelines.R")
source("~/OneDrive - Nexus365/Project ClinCluster/evaluate-integration/evaluate-integration/functions_initial.R")

verbose <- FALSE
dirsave <- "progenitor"

n <- 60
n_size <- 40
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
# image(1:length(col_vector), 1, as.matrix(1:length(col_vector)), col = col_vector)
```
# Read data

```{r readrds}
seu <- readRDS(paste0("rdata/",dirsave,"/seurat_object_cca_integrated.rds"))
```

# Calculate dist

## Compute IDE-based similarity

Composition of clusters

```{r cca cluster}
res <- dbscan::hdbscan(seu@reductions$pca@cell.embeddings[,1:15], minPts = 15)
seu$cluster <- as.factor(res$cluster)
seu$cluster <- as.factor(seu$cluster)

table(seu$Group, seu$cluster)
```

```{r pipeline2-table-cluster-composition}
metadata <- data.frame(label = paste(seu$cluster, seu$Batch, sep = "_"),
                       batch = seu$Batch,
                       ground_truth = seu$Group,
                       V1 = seu@reductions$tsne@cell.embeddings[,1],
                       V2 = seu@reductions$tsne@cell.embeddings[,2], stringsAsFactors = FALSE)

knitr::kable(table(metadata$label, metadata$ground_truth))
```



```{r pipeline2-de-analysis}
system.time( {
  select <- downsampling(metadata = metadata, n_size = n_size, include = TRUE, replace = TRUE) # downsampling
  matrix <- as.matrix(seu@assays$RNA@counts[,select])
  colnames(matrix) <- paste0(colnames(matrix), "_", 1:ncol(matrix))
  keep <- rowSums(matrix > 0.5) > 5
  dge <- edgeR::DGEList(counts = matrix[keep, , drop = F]) # make a edgeR object
  dge <- dge[!grepl("ERCC-", rownames(dge)),] # remove ERCC
  dge <- dge[!grepl("MT-", rownames(dge)),]
  logCPM <- edgeR::cpm(dge, log = TRUE, prior.count = 3)
  
  df <- data.frame(g = metadata$label[select],
                   b = metadata$batch[select], ## batch
                   stringsAsFactors = F) ## label
  df$detrate <- colSums(matrix > 0.5)
  rownames(df) <- colnames(matrix)
  
  N <- length(unique(df$g))
  combinations <- data.frame(g1 = rep(unique(df$g), each = N), g2 = rep(unique(df$g), N), stringsAsFactors = FALSE)
  combinations <- combinations[combinations$g1 != combinations$g2, ]
  combinations$b1 <- df$b[match(combinations$g1, df$g)]
  combinations$b2 <- df$b[match(combinations$g2, df$g)]
  combinations <- combinations[combinations$b1!=combinations$b2,]
  
  idx <- c()
  for(i in 2:nrow(combinations)){
    if(!combinations$g2[i] %in% combinations$g1[1:(i-1)]) {
      idx <- c(idx, i)
    }
  }
  
  combinations <- combinations[c(1,idx),]
  rownames(combinations) <- 1:nrow(combinations)
  
  dist_p <- dist_coef <- matrix(0, nrow = N, ncol = N)
  colnames(dist_p) <- rownames(dist_p) <- unique(df$g)
  colnames(dist_coef) <- rownames(dist_coef) <- unique(df$g)
  
  
  # create progress bar
  if (verbose == TRUE) {
    message(paste0("Generating distance matrix..."))
    pb <- txtProgressBar(min = 0, max = nrow(combinations), style = 3)
    k <- 1
  }
  
  for (i in 1:nrow(combinations)){
    
    if (verbose == TRUE) {
      setTxtProgressBar(pb, k) # progress bar
      k <- k+1
    }
    
    df$tmp <- "bg"
    df$tmp[df$g == combinations$g1[i]] <- "g1"
    df$tmp[df$g == combinations$g2[i]] <- "g2"
    
    design <- model.matrix(~  0 + tmp + b + detrate, data = df) 
    contrast_m <- makeContrasts(contrasts = c("tmpg1-tmpbg", "tmpg2-tmpbg"),
                                levels = design)
    
    fit <- lmFit(logCPM, design)
    group_fit <- contrasts.fit(fit, contrast_m)
    group_fit <- eBayes(group_fit, trend = TRUE, robust = TRUE)
    
    idx1 <- rownames(dist_coef) == combinations$g1[i]
    idx2 <- colnames(dist_coef) == combinations$g2[i]
    dist_coef[idx1, idx2] <- cor(coef(group_fit)[,1], coef(group_fit)[,2])
    dist_p[idx1, idx2] <- cor(-log10(group_fit$p.value)[,1]*sign(coef(group_fit)[,1]), -log10(group_fit$p.value)[,2]*sign(coef(group_fit)[,2]))
  }
  
  if(verbose == TRUE) {
    close(pb) # close progress bar
  }
})
```

```{r}
dist_coef
```


## Final clustering

```{r pipeline2-final clustering}
df <- data.frame(g = metadata$ground_truth[select],
                 b = metadata$batch[select], ## batch
                 c = metadata$label[select], stringsAsFactors = F) ## label
df$combination <- paste0(df$g, "-", df$c)
freq <- table(df$combination)
df$freq <- freq[match(df$combination, names(freq))]
df <- df[order(df$freq, decreasing = TRUE),]
df <- unique(df)

N <- length(unique(df$g))
edges <- data.frame(from = combinations$g1, to = combinations$g2, weight = NA)

tmp <- dist_coef + t(dist_coef)
for(i in 1:nrow(edges)) {
  edges$weight[i] <- tmp[rownames(tmp) == edges$from[i], colnames(tmp) == edges$to[i]]
}

edges$weight[edges$weight < 0] <- 0
edges <- edges[edges$weight > 0, ]
net <- graph_from_data_frame(edges, directed = FALSE)
E(net)$width <- 2^(E(net)$weight * 7)
vg_names <- attr(V(net), "names")
df_cols <- data.frame(group = unique(df$g), col = col_vector[1:length(unique(df$g))], stringsAsFactors = FALSE)
df$col <- df_cols$col[match(df$g, df_cols$group)]

V(net)$color <- df$col[match(vg_names, df$c)]
V(net)$frame.color <- "#777777"
V(net)$size <- 8
V(net)$label.family <- "Helvetica"
plot(net);legend(x=-1.5, y=-1.1, df_cols$group, pch=21, col="#777777", pt.bg=df_cols$col, pt.cex=2, cex=1, bty="n", ncol=1)
```

```{r pipeline2-hirerchical-tree}
tmp <- dist_coef + t(dist_coef)
diag(tmp) <- 1
dist <- as.dist( (1 - tmp) )
plot(hclust(dist))
```

```{r pipeline2-final-clustering-match-names}
hc <- hclust(dist)
hcluster <- cutree(hc, h = 0.45)
evaluation_input <- dist
df_merge <-  data.frame(initial_clusters = names(hcluster),
                        final_clusters = hcluster)

seu$final_cluster <- df_merge$final_cluster[match(metadata$label,df_merge$initial_clusters)]
seu$final_cluster[is.na(seu$final_cluster)] <- metadata$label[is.na(seu$final_cluster)]
knitr::kable(table(seu$final_cluster, seu$Group))
```

# Estimate probability

## Sampling from the presumably same cell population

Assume that the two clusters with the highest similarity are from the same population because they are the mutual nearest neighbors with the highest similarity.

```{r select-positive-control}
pos_control <- c(rownames(dist_coef)[which.max(apply(dist_coef, 1, max))], colnames(dist_coef)[which.max(apply(dist_coef, 2, max))])
pos_control
```

## Calculate the background distribution

```{r further-clustering-within-positive-controls}
## further-clustering-within-positive-controls
system.time({
  seeds <- c(12345, 89465, 10385, 10385, 17396, 34523, 12532, 56734, 12653, 23461)
  idx <- paste(seu$cluster, seu$Batch, sep = "_") %in% pos_control
  
  combinations_all <- c()
  bg_dist_coef_list <- list()
  seu_selected <- seu[,idx]
  pca <- seu@reductions$pca@cell.embeddings[idx, 1:15]
  
  for(itor in 1:10) {
    set.seed(seeds[itor])
    res <- kernlab::specc(pca, centers = 5)
    
    ## Calculate background distribution
    seu_selected$forced_cluster <- res@.Data
    seu$forced_cluster <- seu_selected$forced_cluster[match(colnames(seu), colnames(seu_selected))]
    seu$forced_cluster[is.na(seu$forced_cluster)] <- "bg"
    
    metadata <- data.frame(label = paste0(seu$forced_cluster, "_", seu$Batch),
                           batch = seu$Batch,
                           ground_truth = seu$Group, stringsAsFactors = FALSE)
    table(metadata$label, metadata$ground_truth)
    
    # downsampling
    select <- downsampling(metadata = metadata, n_size = n_size, include = TRUE, 
                           replace = TRUE, lower.cutoff = 15)
    
    # IDER
    matrix <- as.matrix(seu@assays$RNA@counts[,select])
    keep <- rowSums(matrix > 0.5) > 5
    dge <- edgeR::DGEList(counts = matrix[keep, , drop = F]) # make a edgeR object
    dge <- dge[!grepl("ERCC-", rownames(dge)),] # remove ERCC
    dge <- dge[!grepl("MT-", rownames(dge)),]
    
    df <- data.frame(g = metadata$label[select],
                     b = metadata$batch[select], ## batch
                     stringsAsFactors = F) ## label
    df$detrate <- colSums(matrix > 0.5)
    colnames(matrix) <- paste0(colnames(matrix), "_", 1:ncol(matrix))
    rownames(df) <- colnames(matrix)
    
    GROUPS <- unique(df$g)
    N <- length(GROUPS)
    
    combinations <- data.frame(g1 = rep(unique(df$g), each = N), g2 = rep(unique(df$g), N), stringsAsFactors = FALSE)
    combinations <- combinations[combinations$g1 != combinations$g2, ]
    combinations$b1 <- df$b[match(combinations$g1, df$g)]
    combinations$b2 <- df$b[match(combinations$g2, df$g)]
    combinations <- combinations[combinations$b1 != combinations$b2, ]
    idx <- c()
    for(i in 2:nrow(combinations)){
      if(!combinations$g2[i] %in% combinations$g1[1:(i-1)]) {
        idx <- c(idx, i)
      }
    }
    
    combinations <- combinations[c(1,idx),]
    rownames(combinations) <- 1:nrow(combinations)
    combinations <- combinations[!combinations$g1 %in% c("bg_Batch1", "bg_Batch2")  & !combinations$g2 %in% c("bg_Batch1", "bg_Batch2"),]
    combinations$similarity <- NA
    combinations$iteration <- itor
    
    bg_dist_coef <- matrix(0, nrow = N, ncol = N)
    colnames(bg_dist_coef) <- rownames(bg_dist_coef) <- GROUPS
    
    for (i in 1:nrow(combinations)){
      df$tmp <- NA
      df$tmp[df$g %in% c("bg_Batch1", "bg_Batch2")] <- "bg"
      df$tmp[df$g == combinations$g1[i]] <- "g1"
      df$tmp[df$g == combinations$g2[i]] <- "g2"
      
      idx <- which(!is.na(df$tmp))
      design <- model.matrix(~  0 + tmp + b + detrate, data = df[idx, ]) 
      contrast_m <- makeContrasts(contrasts = c("tmpg1-tmpbg", "tmpg2-tmpbg"), levels = design)
      logCPM <- edgeR::cpm(dge[, idx], log = TRUE, prior.count = 3)
      fit <- lmFit(logCPM, design)
      group_fit <- contrasts.fit(fit, contrast_m)
      suppressWarnings(group_fit <- eBayes(group_fit, trend = TRUE, robust = TRUE))
      
      idx1 <- rownames(bg_dist_coef) == combinations$g1[i]
      idx2 <- colnames(bg_dist_coef) == combinations$g2[i]
      combinations$similarity[i] <- bg_dist_coef[idx1, idx2] <- cor(coef(group_fit)[,1], coef(group_fit)[,2])
    }
    combinations_all <- rbind(combinations_all, combinations)
    bg_dist_coef_list[[itor]] <- bg_dist_coef
  }
})
```


## Visualise the background distribution

```{r histogram-background-distribution, fig.width=6}
ggplot(combinations_all, aes(x = similarity)) +
  geom_histogram(fill = alpha(col_vector[2], 0.8),
                 col = "grey40",
                 binwidth = 0.01) + theme_classic() + xlim(0.4, 1)
ggsave(paste0("plots/",dirsave,"/cutoff_selection/bg_similarity_distribution_histogram.pdf"), width = 4, height = 2.5)
```

```{r calculate-empirical-p}
getProbability <- function(x, bg_similarity) { # calculate the probability
  try({
    res <- matrix(NA, ncol = ncol(x), nrow = nrow(x))
    rownames(res) <- rownames(x)
    colnames(res) <- colnames(x)
    len <- length(bg_similarity)
    for(i in 1:nrow(x)){
      for(j in 1:ncol(x)){
        res[i,j] <- sum(bg_similarity > x[i,j]) / len
      }
    }
    return(res)
  })
}

idx1 <- paste0(0:2, "_Batch1")
idx2 <- paste0(0:2, "_Batch2")
```

```{r plot-heatmap-similarity-pvalue}
p_mat <- getProbability(dist_coef[idx1, idx2], combinations_all$similarity)
pheatmap(p_mat, main = "Empirical p-value",
         color = colorRampPalette(c("white", col_vector[8]))(50), 
         filename = paste0("plots/",dirsave,"/cutoff_selection/probability_heatmap.pdf"),
         cellwidth = 25, cellheight = 25, width = 10, height = 10,
         display_numbers = TRUE, cluster_rows = FALSE, cluster_cols = FALSE)
```


```{r plot-heatmap-similarity}
pheatmap(dist_coef[idx1, idx2], main = "Similarity",
         color = colorRampPalette(c("white", col_vector[8]))(50), 
         filename = paste0("plots/",dirsave,"/cutoff_selection/similarity_heatmap.pdf"),
         cellwidth = 25, cellheight = 25, width = 10, height = 10,
         display_numbers = TRUE, cluster_rows = FALSE, cluster_cols = FALSE)
```


# Summary


```{r diffusionmap}
set.seed(12345)
dm <- destiny::DiffusionMap(seu@reductions$pca@cell.embeddings[,1:15])
```

```{r plot-tsne}
idx1 <- paste0(0:2, "_Batch1")
idx2 <- paste0(0:2, "_Batch2")
scores <- diag(dist_coef[idx1, idx2])
names(scores) <- 0:2
seu$similarity <- scores[match(seu$cluster, names(scores))]

scores <- diag(p_mat[idx1, idx2])
names(scores) <- 0:2
seu$pvalue <- scores[match(seu$cluster, names(scores))]
  
df_plot <- data.frame(V1 = dm@eigenvectors[,1],
                      V2 = dm@eigenvectors[,2],
                      label = seu$Group,
                      batch = seu$Batch,
                      cluster = seu$cluster,
                      similarity = seu$similarity,
                      pvalue = seu$pvalue)
p1 <- ggplot(df_plot, aes(x = V1, y = V2, fill = as.factor(label))) + geom_point(col = "grey40", size = 2, shape = 21)  + scale_fill_manual(values = col_vector[1:10]) + theme_classic() + theme(legend.position = "top", axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank()) 
p2 <- ggplot(df_plot, aes(x = V1, y = V2, fill = similarity)) + geom_point(col = "grey40", size = 2, shape = 21)  +  scale_fill_viridis(option = "B") + theme_classic() + theme(legend.position = "top", axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())
p3 <- ggplot(df_plot, aes(x = V1, y = V2, fill = pvalue)) + geom_point(col = "grey40", size = 2, shape = 21)  +  scale_fill_viridis(option = "B", direction = -1) +  
  theme_classic() + theme(legend.position = "top", axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank()) 
plot_grid(p1,p2,p3, ncol = 3)
ggsave(paste0("plots/",dirsave,"/cutoff_selection/results_tSNE.pdf"), width = 14, height = 5, useDingbats = FALSE)
```


# Technical

```{r sessioninfo}
sessionInfo()
```

