---
title: "Test with the simulated dataset"
author: "Zhiyuan"
date: "8/6/2020 (last modified: `r Sys.Date()`)"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Dropbox/CIDER/evaluate-integration")

library(splatter)
library(Seurat)
# BiocManager::install("batchelor")
library(batchelor)
library(scater)
library(scran)
library(SingleCellExperiment)
library(wesanderson)
library(harmony)
library(limma)

library(reticulate)

# devtools::install_github("sctyner/geomnet")
library(igraph)
library(geomnet)
library(RColorBrewer)

source("~/Dropbox/CIDER/evaluate-integration/functions_simulate.R")

qual_col_pals <- brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector <- unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

# Summary

This rmarkdown is aimed to visualise the analysis applied on the Dataset 2.

## Dataset

3 simulated batches

* dataset 1: A & B & C
* dataset 2: B & C & D
* dataset 3: C & D & E

There is something wrong about the scanorama dataset. So I will generate my own dataset.

# Preprocessing

First we generate 3 batches and 4 populations. Only batch 1 type 1&2; batch 2 type 2&3; batch 3 type 3&4 are kept.

```{r splatter-simulation}
set.seed(1000)
params <- newSplatParams()
params <- setParams(params, seed = 1000, 
                    group.prob = rep(1/5, 5), 
                    batchCells = c(2000,2000,2000)
)
sim.groups <- splatSimulate(params, method = "groups",
                            verbose = FALSE)
# remove the populations
sim.groups <- sim.groups[,paste0(sim.groups$Batch, sim.groups$Group) %in%
                           c("Batch1Group1","Batch1Group2","Batch1Group3",
                             "Batch2Group2","Batch2Group3","Batch2Group4",
                             "Batch3Group3","Batch3Group4","Batch3Group5")]
```

```{r preprocess seurat object, fig.width=8, fig.height=4}
logcounts(sim.groups) <- log2(calculateCPM(sim.groups) + 1)
seu <- seurat_preprocess(sim.groups)

df_plot <- data.frame(V1 = seu@reductions$tsne@cell.embeddings[,1],
                      V2 = seu@reductions$tsne@cell.embeddings[,2],
                      label = seu$Group,
                      batch = seu$Batch)

p1 <- ggplot(df_plot, aes(x = V1, y = V2, col = as.factor(label))) + 
  geom_point(alpha = 0.6, size=0.5) + theme_classic() + 
  scale_colour_manual(values = wes_palette("Darjeeling1")) + 
  theme(legend.position = "top",axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),axis.text = element_blank())
p2 <- ggplot(df_plot, aes(x = V1, y = V2, col = as.factor(batch))) + 
  geom_point(alpha = 0.6, size=0.5)  +  theme_classic() + 
  theme(legend.position = "top") + 
  scale_color_manual(values = c(wes_palette("Chevalier1")))+
  theme(axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),
        axis.text = element_blank())
p1 + p2
# ggsave("plots/simulation_nonoverlap/tsne_uncorrected.pdf", width = 8, height = 4, useDingbats = FALSE)
```

# Our pipeline

## Compute IDER-based similarity

First prepare initial clusters

```{r prepare-initial-clusters}
metadata <- data.frame(label = seu$Group,
                       batch = seu$Batch,
                       V1 = seu@reductions$tsne@cell.embeddings[,1],
                       V2 = seu@reductions$tsne@cell.embeddings[,2])
n_size <- 50
select <- downsampling(metadata = metadata, n_size = n_size)
metadata$inicluster <- paste0(metadata$batch, "_", metadata$label)
```

```{r plot-initial-clusters, fig.width=5, fig.height=3}
ggplot(metadata[select,], aes(x = V1, y = V2, col = inicluster)) + geom_point(alpha = 0.6, size=0.5)  +  theme_classic() + theme(legend.position = "right") + scale_color_manual(values = col_vector[1:9])+ theme(axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())+xlab("tSNE_1") + ylab("tSNE_2")
# ggsave("plots/simulation_nonoverlap/tsne_initialcluster_wrappededby_batch.pdf", width = 4, height = 2.6, useDingbats=FALSE)
```

```{r compute-similarity}
matrix <- as.matrix(seu@assays$RNA@counts[,select])
list_dist <- calculate_ider(matrix = matrix, metadata = metadata, verbose = FALSE)
dist_p <- list_dist[[3]]
dist_coef <- list_dist[[1]]
```

```{r hirerchical-tree}
plot(hclust(as.dist(1-(dist_p + t(dist_p)))))
```

Five final clusters were identified




## Network graph

```{r final clustering, fig.width=5, fig.height=5}
df <- data.frame(g = paste(metadata$label, metadata$batch, sep = "_"),
                 b = metadata$batch, ## batch
                 c = metadata$label, stringsAsFactors = F) ## label
df <- unique(df)

N <- length(unique(df$g))
combinations <- data.frame(g1 = rep(unique(df$g), each = N), g2 = rep(unique(df$g), N), stringsAsFactors = FALSE)
combinations <- combinations[combinations$g1 != combinations$g2, ]

idx <- c()
for(i in 2:nrow(combinations)){
  if(!combinations$g2[i] %in% combinations$g1[1:(i-1)]) {
    idx <- c(idx, i)
  }
}

combinations <- combinations[c(1,idx),]
rownames(combinations) <- 1:nrow(combinations)
combinations$b1 <- df$b[match(combinations$g1, df$g)]
combinations$b2 <- df$b[match(combinations$g2, df$g)]
combinations <- combinations[combinations$b1!=combinations$b2,]

edges <- data.frame(from = combinations$g1, to = combinations$g2, weight = NA)

tmp <- dist_p + t(dist_p)
for(i in 1:nrow(edges)) {
  edges$weight[i] <- tmp[rownames(tmp) == edges$from[i], colnames(tmp) == edges$to[i]]
}

edges$weight[edges$weight < 0] <- 0
edges <- edges[edges$weight > 0, ]

net <- graph_from_data_frame(edges, directed = FALSE) # generate graph
E(net)$width <- E(net)$weight * 10
vg_names <- attr(V(net), "names")
vg_names <- substr(vg_names, start = 1, stop = 6)
df_cols <- data.frame(group = paste0("Group",1:5),
                      col = as.character(wes_palette("Darjeeling1")), stringsAsFactors = FALSE)
V(net)$color <- df_cols$col[match(vg_names, df_cols$group)]
V(net)$frame.color <- "#777777"
V(net)$size <- 20
V(net)$label.family <- "Helvetica"

df_width <- data.frame(value = c(0.1, 1, 5, 10 )/10, width = c(0.1, 1, 5, 10 ))

# pdf("plots/simulation_nonoverlap/network_plot.pdf",width = 5, height = 5, useDingbats = FALSE)
plot(net);legend(x=-1.5, y=-1.1, legend = df_cols$group, pch=21,
       col="#777777", pt.bg=df_cols$col, 
       pt.cex=2, cex=.8, bty="n", ncol=1) ; legend(x = 0, y = -1.1, legend = as.factor(df_width$value),
                                                   lwd = df_width$width, bty="n", ncol=1)
# dev.off()
```

# Other pipelines

## CCA correction

```{r cca correction, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
seu.integrated <- cca_integration(seu)

seu.integrated <- FindNeighbors(seu.integrated, dims = 1:15)
seu.integrated <- FindClusters(seu.integrated, resolution = 0.6)

res <- dbscan::hdbscan(seu.integrated@reductions$tsne@cell.embeddings[,1:2], minPts = 25)
seu.integrated$cluster <- as.factor(res$cluster)

df_plot <- data.frame(V1 = seu.integrated@reductions$tsne@cell.embeddings[,1],
                      V2 = seu.integrated@reductions$tsne@cell.embeddings[,2],
                      label = seu.integrated$Group,
                      cluster = seu.integrated$cluster)
p1 <- ggplot(df_plot, aes(x = V1, y = V2, col = as.factor(label))) + geom_point(alpha = 0.6, size=1) + theme_classic() + scale_colour_manual(values = wes_palette("Darjeeling1")) + theme(legend.position = "top",axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())
p2 <- ggplot(df_plot, aes(x = V1, y = V2, col = as.factor(cluster))) + geom_point(alpha = 0.6, size=1)  +  theme_classic() + theme(legend.position = "top") + scale_color_manual(values = c("grey40",wes_palette("Moonrise2")))+ theme(axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())
p1 + p2
# ggsave("plots/simulation_nonoverlap/tsne_cca.pdf", width = 8, height = 4, useDingbats = FALSE)
```

## MNN correction

```{r mnn correction, fig.width=8, fig.height=4}
f.out <- mnn_integration(seu)
f.out <- runTSNE(f.out, dimred="corrected")
res <- dbscan::hdbscan(reducedDim(f.out,"TSNE"), minPts = 75)
f.out$cluster <- as.factor(res$cluster)

df_plot <- data.frame(V1 = reducedDim(f.out, "TSNE")[,1],
                      V2 = reducedDim(f.out, "TSNE")[,2],
                      label = f.out$Group,
                      cluster = f.out$cluster)
p1 <- ggplot(df_plot, aes(x = V1, y = V2, col = as.factor(label))) + geom_point(alpha = 0.6, size=1) + theme_classic() + scale_colour_manual(values = wes_palette("Darjeeling1")) + theme(legend.position = "top",axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())
p2 <- ggplot(df_plot, aes(x = V1, y = V2, col = as.factor(cluster))) + geom_point(alpha = 0.6, size=1)  +  theme_classic() + theme(legend.position = "top") + scale_color_manual(values = c("grey40",wes_palette("Moonrise2")))+ theme(axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())
p1 + p2
# ggsave("plots/simulation_nonoverlap/tsne_mnn.pdf", width = 8, height = 4, useDingbats = FALSE)
```

# Technical

```{r sessioninfo}
sessionInfo()
```

