---
title: "Integration and evaluation of the cell line dataset"
author: "Zhiyuan"
date: "8/5/2020 (last modified: `r Sys.Date()`)"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
# BiocManager::install("batchelor")
# library(batchelor)
library(scater)
# library(scran)
library(SingleCellExperiment)
library(wesanderson)

library(limma)
```

# Summary

* Overall aim: Evaluate the integration results
* Scenario
  - Ground truth is unknown (unannotated datasets; ARI is unapplicable)
  - Want to know if biologically similar cell populations are overlapped after integration
* Approach:
  - Extract overlapping cell populations from each batch
  - Examine if these populations share the same set of marker genes compared to the other cells
  
### Datasets

This is a classic dataset for benchmarking with 

* Three batches
  - 293t 
  - jurkat
  - jurkat_293t_50_50
* Two cell types
  - 293t
  - jurkat


* This cell line dataset was integrated by 3 methods
  - MNN (negative control)
  - Scanorama (positive control)
# Preprocessing

## Read output from scanorama

```{r}
metadata <- read.delim("~/scanorama/data/293t_jurkat/293t_jurkat_corrected_embedding.txt", header = F)
cell_label <- read.delim("~/scanorama/data/cell_labels/293t_jurkat_cluster.txt", header = F)
metadata$label <- cell_label$V1
metadata <- data.frame(metadata)
dim(metadata)
```

## Read matrices and create seurat object

```{r create seurat object}
m1 <- Read10X("~/scanorama/data/293t_jurkat/293t/")
m2 <- Read10X("~/scanorama/data/293t_jurkat/jurkat/")
m3 <- Read10X("~/scanorama/data/293t_jurkat/jurkat_293t_50_50/")

seu <- CreateSeuratObject(counts = cbind(m1, m2, m3), min.features = 600)
rm(m1,m2,m3)
dim(seu)
```

```{r add metadata to seurat object}
seu@meta.data$cell_label <- metadata$label
seu@meta.data$batch <- c(rep("293t", 2885), rep("jurkat", 3257), rep("jurkat_293t_50_50", 3388))
metadata$batch <- seu@meta.data$batch
seu@reductions$scanorama <- metadata[,1:2]
```

```{r preprocess seurat object}
# seu <- NormalizeData(seu, verbose = FALSE)
# seu <- FindVariableFeatures(seu, selection.method = "vst",
#         nfeatures = 2000, verbose = FALSE)
# seu <- ScaleData(seu, verbose = FALSE)
# seu <- RunPCA(seu, npcs = 15, verbose = FALSE)
# seu <- RunUMAP(seu, reduction = "pca", dims = 1:15)
# seu <- RunTSNE(seu, reduction = "pca", dims = 1:15)
```

# Integration - Scanorama results

```{r scanorama output, fig.width=10, fig.height=4}
p1 <- ggplot(metadata, aes(x = V1, y = V2, col = label)) + geom_point(alpha = 0.6, size=0.5) + theme_classic() + scale_colour_manual(values = wes_palette("Chevalier1")[2:1]) + theme(legend.position = "top") + 
  theme(axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())
p2 <- ggplot(metadata, aes(x = V1, y = V2, col = batch)) + geom_point(alpha = 0.6, size=0.5) + theme_classic() + scale_colour_manual(values = wes_palette("Moonrise2"))  + theme(legend.position = "top")+ 
  theme(axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())
# p1 + p2
```


# Clustering

Apply DBSCAN on the corrected low-dimensional space


Scanorama performs well on this 

```{r Scanorama clustering, fig.width=12, fig.height=5}
df_scan <- metadata[,1:2] ## scanorama
df_scan <- as.data.frame(df_scan)
res <- dbscan::hdbscan(df_scan[,1:2], minPts = 75)
df_scan$cell_type <- metadata$label
df_scan$cluster <- as.factor(res$cluster)
df_scan$batch <- as.factor(c(rep("293t", 2885), rep("jurkat", 3257), rep("jurkat_293t_50_50", 3388)))

# p1 <- ggplot(df_scan, aes(x = V1, y = V2, col = cluster)) + geom_point(alpha = 0.6)  +  theme_void() + theme(legend.position = "top")
# p2 <- ggplot(df_scan, aes(x = V1, y = V2, col = cell_type)) + geom_point(alpha = 0.6) + theme_void() + theme(legend.position = "top")
# p3 <- ggplot(df_scan, aes(x = V1, y = V2, col = batch)) + geom_point(alpha = 0.6) +     theme_void() + theme(legend.position = "top")
# 
# p_list <- list()
# p_list[[1]] <- p1
# p_list[[2]] <- p2
# p_list[[3]] <- p3

p3 <- ggplot(df_scan, aes(x = V1, y = V2, col = cluster)) + geom_point(alpha = 0.6, size=0.5)  +  theme_classic() + theme(legend.position = "top") + scale_color_manual(values = c("grey40",wes_palette("Darjeeling1")[2:3]))+ 
  theme(axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())

p2+p1+p3
```

```{r}
# ggsave(plot = p2+p1+p3, filename = "plots/cell_line_dataset/tSNE_scanorama_all.pdf", width = 11, height = 4, useDingbats = FALSE)
```


```{r cluster results dataframe}
cluster_df <- data.frame(scanorama_cluster = df_scan$cluster, 
                         batch = seu$batch)
```

# DE analysis

Composition of clusters

```{r scanorama cluster composition}
knitr::kable(table(cluster_df$scanorama_cluster, cluster_df$batch))
```

To get a more evenly distributed composition for DE analysis, we choose 50 cells for each group (cluster + batch).

```{r scanorama-select-50-cells-per-category, fig.width=10, fig.height=3}
n_size <- 50
cluster <- c(1:2)
tech <- c("293t", "jurkat","jurkat_293t_50_50")
select <- c()
for(i in cluster){
  for(j in tech) {
    idx <- which(cluster_df$scanorama_cluster %in% i & cluster_df$batch %in% j)
    if(length(idx) >= 50){
      set.seed(12345)
      select <- c(select, sample(idx, size = n_size, replace = F))
    }
  }
}

ggplot(df_scan[select,], aes(x = V1, y = V2, col = as.factor(cluster))) + geom_point() + theme_classic() + facet_wrap(~batch) + scale_colour_manual(values = wes_palette("Darjeeling1")[2:3]) + 
  theme(legend.position = "top",axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank())
# ggsave("plots/cell_line_dataset/scanorama_cluster_wraped_batch.pdf", width = 11, height = 4, useDingbats=FALSE) # https://stackoverflow.com/questions/39707542/geom-point-shape-discrepancy-between-pdf-and-png-output?noredirect=1&lq=1
```

```{r scanorama-de-analysis}
matrix <- as.matrix(seu@assays$RNA@counts[,select])
keep <- rowSums(matrix > 0.5) > 5 
dge <- edgeR::DGEList(counts = matrix[keep,,drop=F]) # make a edgeR object
dge <- dge[!grepl("ERCC-", rownames(dge)),]
rm(matrix, keep)

df <- data.frame(g = paste(cluster_df$scanorama_cluster, cluster_df$batch, sep = "_")[select],
                 b = cluster_df$batch[select], ## batch
                 c = cluster_df$scanorama_cluster[select], stringsAsFactors = F)
df$detrate <- colSums(seu@assays$RNA@counts[,select] > 0.5)
rownames(df) <- colnames(seu)[select]

combinations <- data.frame(g1 = rep(unique(df$g), each = 4), g2 = rep(unique(df$g), 4), stringsAsFactors = FALSE)
combinations <- combinations[combinations$g1 != combinations$g2, ]

dist_p <- dist_t <- dist_coef <- matrix(0, nrow = 4, ncol = 4)
colnames(dist_p) <- rownames(dist_p) <- unique(df$g)
colnames(dist_t) <- rownames(dist_t) <- unique(df$g)
colnames(dist_coef) <- rownames(dist_coef) <- unique(df$g)

for(i in 1:nrow(combinations)){
  df$tmp <- "bg"
  df$tmp[df$g == combinations$g1[i]] <- "g1"
  df$tmp[df$g == combinations$g2[i]] <- "g2"

  ## by group
  design <- model.matrix(~  0 + tmp + b + detrate, data = df)  # Use 0 because we do not need intercept for this linear model
  v <- voom(dge, design, plot = F) 
  fit_g <- lmFit(v, design) 
  groups <- paste0("tmp", unique(df$tmp))
  groups <-  groups[groups != "tmpbg"]
  perm_groups <- data.frame(g1 = groups,
                            g2 = "tmpbg", stringsAsFactors = F)
  perm_groups <- perm_groups[perm_groups$g1 != perm_groups$g2,]
  perm_groups$pair <- paste0(perm_groups$g1, "-", perm_groups$g2)
  contrast_m <- makeContrasts(contrasts = perm_groups$pair, 
                              levels = design)
  group_fit <- contrasts.fit(fit_g, contrast_m) # Compute Contrasts from Linear Model Fit
  group_fit <- eBayes(group_fit)
  
  dist_coef[rownames(dist_coef) == combinations$g1[i], colnames(dist_coef) == combinations$g2[i]] <- cor(coef(group_fit)[,1], coef(group_fit)[,2])
  dist_t[rownames(dist_t) == combinations$g1[i], colnames(dist_t) == combinations$g2[i]] <- cor(group_fit$t[,1], group_fit$t[,2])
  dist_p[rownames(dist_p) == combinations$g1[i], colnames(dist_p) == combinations$g2[i]] <- cor(-log10(group_fit$p.value)[,1]*sign(coef(group_fit)[,1]), -log10(group_fit$p.value)[,2]*sign(coef(group_fit)[,2]))
}
```

```{r}
# similarity
dist_p
```

```{r}
dist_t
```

## barplot

```{r barplot, fig.height=3, fig.width=6.5}
dist_df <- data.frame(pair = c("g1-g1`", "g2-g2`", "g1-g2", "g1-g2`","g1`-g2","g1`-g2`"),
                      similarity = c(dist_t[1,2],
                                     dist_t[3,4],
                                     dist_t[1,3],
                                     dist_t[1,4],
                                     dist_t[2,3],
                                     dist_t[2,4]),
                      type = c(rep("Identical cell types",2),
                               rep("Different cell types", 4)))

dist_df$pair <- factor(dist_df$pair, levels = dist_df$pair)
ggplot(dist_df, aes(x = pair, y = similarity, fill = type)) + geom_bar(stat ="identity") + ylab("Similarity measured by ClinCluster") + scale_fill_manual(values = wes_palette("Royal1")) + theme_classic()
# ggsave("plots/cell_line_dataset/similarity_barplots.pdf", width = 5.5, height = 3)

```


```{r}
plot(hclust(as.dist(1-dist_t)))
```


## Two examples - scatter plots

```{r}
# example 1
i <- 1

df$tmp <- "bg"
df$tmp[df$g == combinations$g1[i]] <- "g1"
df$tmp[df$g == combinations$g2[i]] <- "g2"

## by group
design <- model.matrix(~  0 + tmp + b + detrate, data = df)  # Use 0 because we do not need intercept for this linear model
v <- voom(dge, design, plot = F) 
fit_g <- lmFit(v, design) 
groups <- paste0("tmp", unique(df$tmp))
groups <-  groups[groups != "tmpbg"]
perm_groups <- data.frame(g1 = groups,
                          g2 = "tmpbg", stringsAsFactors = F)
perm_groups <- perm_groups[perm_groups$g1 != perm_groups$g2,]
perm_groups$pair <- paste0(perm_groups$g1, "-", perm_groups$g2)
contrast_m <- makeContrasts(contrasts = perm_groups$pair, 
                            levels = design)
group_fit <- contrasts.fit(fit_g, contrast_m) # Compute Contrasts from Linear Model Fit
group_fit <- eBayes(group_fit)

df_plot <- data.frame(d1 = group_fit$t[,1], d1prime = group_fit$t[,2])
p1 <- ggplot(df_plot, aes(x = d1, y = d1prime)) + geom_point(alpha = 0.2) + theme_classic()  + xlim(-49, 44) + ylim(-43, 30)
 
# example 2
i <- 5

df$tmp <- "bg"
df$tmp[df$g == combinations$g1[i]] <- "g1"
df$tmp[df$g == combinations$g2[i]] <- "g2"

## by group
design <- model.matrix(~  0 + tmp + b + detrate, data = df)  # Use 0 because we do not need intercept for this linear model
v <- voom(dge, design, plot = F) 
fit_g <- lmFit(v, design) 
groups <- paste0("tmp", unique(df$tmp))
groups <-  groups[groups != "tmpbg"]
perm_groups <- data.frame(g1 = groups,
                          g2 = "tmpbg", stringsAsFactors = F)
perm_groups <- perm_groups[perm_groups$g1 != perm_groups$g2,]
perm_groups$pair <- paste0(perm_groups$g1, "-", perm_groups$g2)
contrast_m <- makeContrasts(contrasts = perm_groups$pair, 
                            levels = design)
group_fit <- contrasts.fit(fit_g, contrast_m) # Compute Contrasts from Linear Model Fit
group_fit <- eBayes(group_fit)

df_plot <- data.frame(d1 = group_fit$t[,1], d1prime = group_fit$t[,2])
p2 <- ggplot(df_plot, aes(x = d1, y = d1prime)) + geom_point(alpha = 0.2) + theme_classic() + xlim(-49, 44) + ylim(-43, 30)
```

```{r scatterplot, fig.width=8, fig.height=4}
p1 + p2
# ggsave(plot = p1+p2, filename = "plots/cell_line_dataset/scatterplot_d1_d1prime.pdf",width = 8, height = 4, useDingbats=FALSE)
```


# Technical

```{r sessionInfo}
sessionInfo()
```



